Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge for Model Modification

### 2024-08-07 paper/main.tex
Updates

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
	development of numerous automated algorithms for constructing DAGs from
	data. Despite these advancements, the application of these algorithms
	in applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use automated algorithms arises due to several
	practical issues, such as algorithms often make obvious mistakes making
	it difficult to trust, they output the Markov Equivalence Class (MEC)
	instead of a DAG (which is often required for further analysis), and
	difficulties in choosing the appropriate algorithm and hyperparameters
	for a given dataset. Even when using these algorithms, researchers must
	rely on their expert knowledge to make modifications and orient edges
	of the learned model. Given this reliance on manual model modification,
	in this paper, we present a method inspired by \emph{specification
	search} in \emph{Structural Equation Models} (SEMs) to assist
	researchers in iteratively modifying their models to improve their fit
	to data. Specifically, we utilize implied Conditional Independence (CI)
	based model testing combined with a novel measure of partial
	association for mixed data to rank violations of implied CIs of the
	model in data. Using this ranking users can choose to make
	modifications to their models by either removing edges or adding edges
	between pair of variables while using their expert knowledge to decide
	the orientation of the edges. The ranking allows users to prioritize
	the worst offending violations first. Through empirical analysis, when
	applied to learning the model from scratch, we demonstrate that taking
	a greedy approach (i.e., fixing the worst violation first) this
	approach can get comparable performance to automated algorithms if the
	users are able to decide the edge direction correctly with an accuracy
	higher than $ 0.4 $. (TODO: Double check this value). We also show that
	a Large Language Model (LLM) can also be used for the purpose of
	deciding the edge orientation in the absence of an expert.
\end{abstract}

>>>>

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
	development of numerous automated algorithms for learning DAGs from
	data. Despite these advancements, the application of these algorithms
	in applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use these algorithms arises due to several
	practical issues, such as algorithms often make obvious mistakes making
	it difficult to trust, they output the Markov Equivalence Class (MEC)
	instead of a DAG (which is often required for further analysis), and
	difficulties in choosing the appropriate algorithm and hyperparameters
	for a given dataset. Even when using these algorithms, researchers must
	rely on their expert knowledge to make modifications and orient edges
	of the learned model. Given this reliance on manual model modification,
	in this paper, we present a method inspired by \emph{specification
	search} in \emph{Structural Equation Models} (SEMs) to assist
	researchers in iteratively modifying their models to improve their fit
	to data. Specifically, we utilize implied Conditional Independence (CI)
	based model testing combined with a novel measure of partial
	association for mixed data to rank violations of implied CIs of the
	model in data. Using this ranking users can choose to make
	modifications to their models by either removing edges or adding edges
	between pair of variables while using their expert knowledge to decide
	the orientation of the edges. The ranking allows users to prioritize
	the worst offending violations first. Through empirical analysis, when
	applied to learning the model from scratch, we demonstrate that taking
	a greedy approach (i.e., fixing the worst violation first) this
	approach can get comparable performance to automated algorithms if the
	users are able to decide the edge direction correctly with an accuracy
	higher than $ 0.4 $. (TODO: Double check this value). We also show that
	a Large Language Model (LLM) can also be used for the purpose of
	deciding the edge orientation in the absence of an expert.
\end{abstract}
